REDO Simulating Dwarf Galaxy Evolution with Individual Star Feedback and Chemical Evolution I: Methods and Test Case

### 2017-11-29 main.tex
up to 5.3

Unfortunately, these high resolution simulations are not (yet) run for the $\sim$ Gyr timescales required to make direct comparisons to the observed stellar and gas phase metallicities of comparable dwarf galaxies at z = 0. However, we can compare to a key observable, the ejection / retention fraction of metals within stars and the galaxy's ISM compared to what would be expected from stellar evolution models based on the galaxy's star formation history. This can be done readily with Milky Way dSph's, whose stars seem to retain very little of the expected metals produced, on order of a few percent or less depending on the galaxy and the species \citep{Kirby2011-metals}. However, environmental effects, namely ram pressure and tidal stripping, complicate the understanding of how these metals were removed from the galaxy. Leo P, the dwarf galaxy we approximate in our initial conditions, is extremely valuable as a gaseous, star forming low mass dwarf galaxy close enough to the Milky Way to conduct this experiment and with an observable \ion{H}{2} region, necessary for determining gas phase abundances. Leo P retains $\sim$ 5\% $\pm$ 2\% of its metals, $\sim$ 1\% in stars and the rest in ISM gas \citep{McQuinn2015}. As discussed in Section~\ref{sec:chemical evolution}, roughly \textbf{95\%} of the tracked metals produced throughout the simulation time no longer reside within the galaxy's disk. At the end of our simulation, we agree with the observations; however, this is an evolving quantity that also depends on how much (if any) subsequent re-accretion of these metals occurs. Although more than half of these metals are expected to be re-accreted \citep{Angles-Alcazar2017}, it is still possible that, at any moment, most of the metals that have been produced lie outside the galaxy disk. We will examine the long term, and species-by-species, evolution of the metal retention fraction in future work.

It is interesting to consider where these metals lie outside of the galaxy's disk, in the CGM or ejected into the IGM, and their eventual fate. While this cannot be observationally determined for dwarf galaxies at this mass, cosmological simulations of a $M_{\rm vir} = 10^{10}$ M$_{\odot}$ galaxy show that, by redshift zero $\sim$ 60\% of ejected gas is recycled over time, while 40\% is ejected from the galaxy's halo \citep{Angles-Alcazar2017}. In our case, no metals have been ejected outside the virial radius; this is possibly not surprising however given that metals ejected outside the halo would need to travel with an average velocity of 150 km s$^{-1}$ for the entirety of our simulation time. \textbf{32\%} of all metals lie within 1 kpc of the center of our galaxy, \textbf{75\%} within 0.25 $R_{\rm vir}$ (or 6.85 kpc), and only \textbf{10\%} outside $R_{\rm vir}$. Over time, the re-accretion or final ejection of this gas is directly relevant to the chemical evolution of low mass dwarf galaxies. Re-cycling of metal enriched gas, could be a significant source of long-term chemical evolution in low mass galaxies, particularly if a majority of metals ejected from the disk (itself nearly all the metals produced by the galaxy) returns. Unfortunately, we cannot yet speak to how this gas evolves over time, but will investigate this in future work with our feedback model.

>>>>

%mm Unfortunately, these high resolution simulations are not (yet) run
    The simulations presented here have not yet been run
for the 
%mm $\sim$ Gyr 
    gigayear
timescales required to make direct comparisons to the observed stellar and gas phase metallicities of comparable dwarf galaxies at $z = 0$. However, we can compare to a key observable, the 
%mm ejection / 
retention fraction of metals within stars and the galaxy's ISM compared to what would be expected from stellar evolution models based on the galaxy's star formation history. This can be done readily with Milky Way dSph's, whose stars seem to retain very little of the expected metal production: on order of a few percent or less depending on the galaxy and the species \citep{Kirby2011-metals}. However, environmental effects, namely ram pressure and tidal stripping, complicate the understanding of how these metals were removed from the galaxy.

Leo P, the dwarf galaxy we approximate in our initial conditions, is extremely valuable as a gaseous, star forming low-mass dwarf galaxy, with an observable \ion{H}{2} region, necessary for determining gas phase abundances, and close enough to the Milky Way to conduct this experiment. Leo P retains $\sim$ 5\% $\pm$ 2\% of its metals, $\sim$ 1\% in stars and the rest in ISM gas \citep{McQuinn2015}. As discussed in Section~\ref{sec:chemical evolution}, roughly \textbf{95\%} of the tracked metals produced during the simulation no longer reside within the galaxy's disk. At the end of our simulation, we agree with the observations; however, this is an evolving quantity that also depends on how much (if any) subsequent re-accretion of these metals occurs. Although more than half of these metals are expected to 
%mm
   eventually
be re-accreted \citep{Angles-Alcazar2017}, it is still possible that, at %mm any moment, 
   the early times we have simulated,
most of the metals that have been produced 
%mm lie 
    remain
outside the galaxy disk. We will examine the long term, and species-by-species, evolution of the metal retention fraction in future work.

It is interesting to consider 
%mm where these metals lie outside of the galaxy's disk, 
    whether ejected metals in our model reside 
in the CGM or have been ejected into the IGM, and their eventual fate. While this cannot be observationally determined for dwarf galaxies at this mass, cosmological simulations of a $M_{\rm vir} = 10^{10}$ M$_{\odot}$ galaxy show that, by redshift zero $\sim$ 60\% of ejected gas is recycled, while 40\% is ejected from the galaxy's halo \citep{Angles-Alcazar2017}. In our case, no metals have been ejected beyond the virial radius; this is not surprising however, given that metals ejected beyond the halo would need to travel with an average velocity of 150 km s$^{-1}$ for the entirety of our simulation time.  We find \textbf{32\%} of all metals lie within 1 kpc of the center of our galaxy, \textbf{75\%} within 0.25 $R_{\rm vir}$ (or 6.85 kpc), and only \textbf{10\%} outside $R_{\rm vir}$. Over time, the re-accretion or final ejection of this gas is directly relevant to the chemical evolution of low mass dwarf galaxies. Recycling of metal enriched gas could be a significant 
%mm source 
    driver
of long-term chemical evolution in low mass galaxies, particularly if a majority of metals ejected from the disk (itself nearly all the metals produced by the galaxy) returns. Unfortunately, we cannot yet speak to how this gas evolves over time, but will investigate this in future work with our feedback model.
